Shrinking Characteristics of Precision Matrix Estimators: An Illustration via Regression

### 2019-05-18 Manuscript.tex
Deploy MGallow/oral_manuscript to github.com/MGallow/oral_manuscript.git:gh-pages

\title{Shrinking Characteristics of Precision Matrix Estimators: An Illustration via Regression}
\author{Matt Galloway}
\date{2019-05-01}

@@

In their 2017 paper, titled \emph{Shrinking Characteristics of Precision Matrix Estimators}, Aaron Molstad, Ph.D.~and Professor Adam Rothman outline a framework to shrink a \emph{characteristic} of a precision matrix. This concept, inspired by others like \citet{cai2011direct}, \citet{fan2012road}, and \citet{mai2012direct}, exploits the fact that in many predictive models estimation of the precision matrix is only necessary through its product with another feature, such as a mean vector. The example they offer in \citet{molstad2017shrinking} is in the context of Fisher's linear discriminant analysis model. If a response vector \(Y\) is categorical such that \(Y\) can take values in \(\left\{1, ..., J\right\}\), then the linear discrimnant analysis model assumes that the design matrix \(X\) conditional on the response vector \(Y\) is normally distributed:

@@

Figure \ref{fig:scpmesim2} displays the model error for each method by dimension of the design matrix. Here we took the sample size equal to \(n = 100\) and the response matrix dimension \(r = 10\) with each data generating procedure replicated a total of 20 times. Note \texttt{OLS} and \texttt{oracleS} are not shown due to extremely poor performance. We find that in high dimensional settings, \texttt{shrinkBO} and \texttt{shrinkB} perform increasingly well relative to the others as the predictor dimension increases. \texttt{shrinkBO} performed the best. Interestingly, when \(n > p\) \texttt{shrinkBO} is still one of the best-performing estimators - though worse than both \texttt{lasso} and \texttt{ridge} - but the performance of \texttt{shrinkB} decreases drastically (plot not shown).

@@

\caption{The oracle precision matrices were tri-diagonal with variable dimension (p) and the data was generated with sample size n = 100 and response vector dimension r = 10. The model errors (ME) for each estimator with variable dimension of the design matrix are plotted. shrinkBO and shrinkB were the two best-performing estimators closely follow by the ridge and lasso regression estimators.}\label{fig:scpmesim2}
\end{figure}

@@

\begin{equation}
\begin{split}
  l\left( \Omega_{y | x}, \Omega_{x}, \Sigma_{xy} | Y, X \right) &= constant + \frac{n}{2}\log\left| \Omega_{y | x} \right| - \frac{1}{2}\sum_{i = 1}^{n} tr\left[ \left( Y_{i} - \Sigma_{xy}'\Omega_{x} X_{i} \right)\left( Y_{i} - \Sigma_{xy}'\Omega_{x} X_{i} \right)'\Omega_{y | x} \right] \\
  &+\frac{n}{2}\log\left| \Omega \right| - \frac{1}{2}\sum_{i = 1}^{n}tr\left( X_{i}X_{i}'\Omega_{x} \right) \\
  &= constant + \frac{n}{2}\log\left| \Omega_{y | x} \right| - \frac{1}{2}tr\left[ \left( \mathbb{Y} - \mathbb{X}\Omega_{x}\Sigma_{xy} \right)'\left( \mathbb{Y} - \mathbb{X}\Omega_{x}\Sigma_{y | x} \right)\Omega_{y | x} \right] \\
  &+ \frac{n}{2}\log\left| \Omega_{x} \right| - \frac{n}{2}tr\left( S_{x}\Omega_{x} \right)
\end{split}
\label{eq:partialjointloglik}\notag
\end{equation}

>>>>

\title{Shrinking Characteristics of Precision Matrix Estimators: An Illustration via Regression}
\author{Matt Galloway}
\date{2019-05-18}

@@

In their 2017 paper, titled \emph{Shrinking Characteristics of Precision Matrix Estimators}, Aaron Molstad, Ph.D.~and Professor Adam Rothman outline a framework to shrink a \emph{characteristic} of a precision matrix. This concept, inspired by others like \citet{cai2011direct}, \citet{fan2012road}, and \citet{mai2012direct}, exploits the fact that in many predictive models estimation of the precision matrix is only necessary through its product with another feature, such as a mean vector. The example they offer in \citet{molstad2017shrinking} is in the context of Fisher's linear discriminant analysis model. If a response vector \(Y\) is categorical such that \(Y\) can take values in \(\left\{1, ..., J\right\}\), then the linear discrimnant analysis model assumes that the predictor matrix \(X\) conditional on the response vector \(Y\) is normally distributed:

@@

Figure \ref{fig:scpmesim2} displays the model error for each method by dimension of the predictor matrix. Here we took the sample size equal to \(n = 100\) and the response matrix dimension \(r = 10\) with each data generating procedure replicated a total of 20 times. Note \texttt{OLS} and \texttt{oracleS} are not shown due to extremely poor performance. We find that in high dimensional settings, \texttt{shrinkBO} and \texttt{shrinkB} perform increasingly well relative to the others as the predictor dimension increases. \texttt{shrinkBO} performed the best. Interestingly, when \(n > p\) \texttt{shrinkBO} is still one of the best-performing estimators - though worse than both \texttt{lasso} and \texttt{ridge} - but the performance of \texttt{shrinkB} decreases drastically (plot not shown).

@@

\caption{The oracle precision matrices were tri-diagonal with variable dimension (p) and the data was generated with sample size n = 100 and response vector dimension r = 10. The model errors (ME) for each estimator with variable dimension of the predictor matrix are plotted. shrinkBO and shrinkB were the two best-performing estimators closely follow by the ridge and lasso regression estimators.}\label{fig:scpmesim2}
\end{figure}

@@

\begin{equation}
\begin{split}
  l\left( \Omega_{y | x}, \Omega_{x}, \Sigma_{xy} | Y, X \right) &= constant + \frac{n}{2}\log\left| \Omega_{y | x} \right| - \frac{1}{2}\sum_{i = 1}^{n} tr\left[ \left( Y_{i} - \Sigma_{xy}'\Omega_{x} X_{i} \right)\left( Y_{i} - \Sigma_{xy}'\Omega_{x} X_{i} \right)'\Omega_{y | x} \right] \\
  &+\frac{n}{2}\log\left| \Omega \right| - \frac{1}{2}\sum_{i = 1}^{n}tr\left( X_{i}X_{i}'\Omega_{x} \right) \\
  &= constant + \frac{n}{2}\log\left| \Omega_{y | x} \right| - \frac{1}{2}tr\left[ \left( \mathbb{Y} - \mathbb{X}\Omega_{x}\Sigma_{xy} \right)'\left( \mathbb{Y} - \mathbb{X}\Omega_{x}\Sigma_{xy} \right)\Omega_{y | x} \right] \\
  &+ \frac{n}{2}\log\left| \Omega_{x} \right| - \frac{n}{2}tr\left( S_{x}\Omega_{x} \right)
\end{split}
\label{eq:partialjointloglik}\notag
\end{equation}
